---
title: "Shouldn't Building Age be Considered when Evaluating Apartment Building Safety^[Code and Data are available at:https://github.com/NnnnninaW/Toronto-Apartment-Building-Evaluation.git] "
author: "Sijian Wu"
date: "6 Feburary 2022"
output: pdf_document
abstract: "Apartment building safety is essential for both landlords and tenants. Dataset of Apartment Building Evaluation from Open Data Toronto to analyze the current status of building safety evaluation which does not appear to have capabilities of detecting severe potential risks like construction damage, hazards resistance, and housing age problem. Through aging, any facility may experience progressive growth of risk, due to deterioration of building materials or other natural physical forces.  To provoke safety awareness for everyone, expanding the diversity and coverage of building safety evaluation is necessary as next step for our city."
bibliography: reference.bib
---

# Table of Contents
### 1  Introduction\

--------------------------------------------------------------------------------------------------2


### 2  Data
#### 2.1 Data Source\

------------------------------------------------------------------------------------------------3

#### 2.2 Methodology and Data Collection\

------------------------------------------------------------------------------------------------3

#### 2.3 Data Characteristics\

------------------------------------------------------------------------------------------------5

### 3 Limitations and Conclusion\

------------------------------------------------------------------------------------------------9

### 4 Reference

------------------------------------------------------------------------------------------------10


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, fig.pos = "H")

```


```{r, include=FALSE, message=FALSE}
library(opendatatoronto)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(ggplot2)
library(modelsummary)
library(estimatr)


data <- read.csv("/Users/nina/Desktop/304 Paper1/Toronto-Apartment-Building-Evaluation/Inputs/DATA/Apartment Building Evaluation.csv")
data
```

```{r, include=FALSE}
#Data Preparation
#Checking some basic characteristics of the original data like type of observations and class, to see if cleaning process is needed for my analysis
glimpse(data)
unique(data$YEAR_REGISTERED)
unique(data$YEAR_EVALUATED)
unique(data$YEAR_BUILT)
unique(data$PROPERTY_TYPE)
sapply(data, class)
unique(data$WARDNAME)

#Changing the YEAR_BUILT,YEAR_EVALUATED, CONFIRMED_STOREYS, CONFIRMED_UNITS into numerical variables for further cleaning
data$YEAR_BUILT <- as.numeric(data$YEAR_BUILT)
class(data$YEAR_BUILT)
data$YEAR_EVALUATED <- as.numeric(data$YEAR_EVALUATED)
class(data$YEAR_EVALUATED)

data$CONFIRMED_STOREYS <- as.numeric(data$CONFIRMED_STOREYS)
class(data$CONFIRMED_STOREYS)
data$CONFIRMED_UNITS <- as.numeric(data$CONFIRMED_UNITS)
class(data$CONFIRMED_UNITS)

colnames(data)
```

```{r, include=FALSE}
#Data Cleaning Section
data_cleaned <- data %>%
   #Convert variables involved in calculation of the final score into numerical variables for analysis
  mutate_at(c(13:35,37,38), as.numeric)%>% 
 #Removing observations that has not applicable information of time for evaluation and time built.
  filter(!is.na(YEAR_EVALUATED), !is.na(YEAR_BUILT))%>%
  #According  to the RentSafeTO, this program is fore Apartment Building with three or more storey or 10 or more units to comply with building maintenance standards. 
  filter(CONFIRMED_STOREYS >= 3 || CONFIRMED_UNITS >= 10)%>%
  #Creating a New Variable to show the building age when encountering the evaluation.
  mutate(Building_Age_at_Evaluation = YEAR_EVALUATED - YEAR_BUILT)%>%
  #Create a single measurement for location as distance to downtown using CN Tower (latitude = 43.642567,-79.387054) as the benchmark, transform degree to km(*110).
  mutate(Distance_to_DT = sqrt((LATITUDE-43.642567)^2+(LONGITUDE+79.387054)^2)*110) %>%
  #Create a new variable that separate observations into groups by the score they have under evaluation, the group is defined by RentSafeTO's criteria of when to conduct the next evaluation
  mutate(Score_Group = 
           case_when(SCORE < 65 ~ "65% or Less",
                     SCORE < 85 ~ "66%-85%",
                     SCORE < 100  ~ "85% or Higher"
                     ))%>%
  #Create a new variable that separate observations into groups by the building age when conducting the evaluation using mutated variable "Building_Age_at_Evaluation"
    mutate(Building_Age_Group = 
           case_when(Building_Age_at_Evaluation < 30  ~ "0-30yr",
                     Building_Age_at_Evaluation < 60 ~ "30-60yr",
                     Building_Age_at_Evaluation < 90 ~ "60-90yr",
                     Building_Age_at_Evaluation < 120 ~ "90-120yr",
                     Building_Age_at_Evaluation < 300 ~ "120+yr"
                     ))%>%
  select(c(6,13,15:35,41,42,43,44))
  
  
data_cleaned
sapply(data_cleaned, class)
summary(data_cleaned$Building_Age_at_Evaluation)
colnames(data_cleaned)
```

\newpage
# Introduction
Apartment building safety is essential for both landlords and tenants. 
The City of Toronto ensures that tenants live in safe, well-maintained buildings through RentSafeTO: Apartment Building Standards, a bylaw enforcement program established in 2017. The program covers buildings that are three or more storeys and ten or more units[@data]. For Housing Safety Management, Housing safety awareness is crucial. Initiatives like evaluations along with effective audits and penalties are essential for provoking safety awareness for everyone. Moreover, it is necessary for programs operated by RentSafeTO to share informative results to the public. 

However, there is no guarantee that the current evaluation criteria could capture all kinds of risks. Risk  could arise because of a specific hazard—an act, event, or phenomenon—posing potential harm to people or activities or things. Fire, earthquake, wind storms, flooding, toxic and allergenic materials, and terrorist attack are examples of hazards associated with buildings[@risk]. It is reasonable for building evaluation focusing on the aspect of risk resistance which is closely related to the building age problem. But, throughout my analysis, the evaluation program does not appear to have capabilities of detecting severe potential risks like construction damage, hazards resistance, and housing age problem. These factors are closely related to individuals not only from the perspective of real estate investment but also personal and public safety.

The remainder sections would show that it is worth paying attention of building age problem in Toronto when evaluating apartment safety since the City of Toronto has experienced rapid urbanization like Shanghai[@shanghai]. Section 2 would present the process of analysis. Section 2.1 would introduce the background and overview of collected data. Section 2.2 would reveal the methods and data cleaning process. Section 2.3 would present results and findings in the analysis. Section 3 concludes our analysis with discussion of limitations. Last Section includes all software and literature materials used for this analysis.

\newpage
# 2 Data
## 2.1 Data Source
To analyze the current status of building safety evaluation in Toronto, I utilized the Apartment Building Evaluation data[@data] from Open Data Toronto [@open]. This dataset contains building evaluation scores for buildings registered with RentSafeTO with individual scores for each criteria. The dataset also contains registered information of when the building is built and what property type it belongs to. However, this dataset does not provide historical scores for each observation which means that each score represents the most recent conducted evaluation. The dataset was last updated on February 5th,2022. Here is a summary of what should be expected as the average level for our analysis in terms of overall score and number of areas being evaluated in Table 1.


```{r, echo=FALSE, message=FALSE }
#Here is a overall summary presenting the mean and standard deviation of evaluation scores and number of items evaluated for each observation
Mean <- lapply(data_cleaned[c(2,3)], mean)
Standard_Deviation <- lapply(data_cleaned[c(2,3)], sd)
table_1 <- cbind(Mean, Standard_Deviation)
knitr::kable(table_1, digits = 4, caption = "Summary of Score and Number of Areas Evaluated")

```
## 2.2 Methodology and Data Collection

Under Apartment Building Evaluation by RentSafeTO, individual criteria are inspected and assigned a score from 1 to 5, with 1 being the lowest and 5 being the highest. If an item is not applicable to the building at the time of evaluation, the score will show as blank in the dataset[@data]. It is reasonable for a building to have a few unavailable examinations due to different infrastructures. Overall, we do have large enough sample size for analysis. The overall score is then determined by dividing the sum of all assigned scores of each item by 5 multiplying the number of unique items reviewed[@data].\

I started my exploratory analysis by using R [@r], dplyr[@dplyr], tidyverse[@tidyverse]. Tables and graphs are generated using kableExtra[@kable], ggplot2[@ggplot2], modelsummary[@modelsummary], and estimatr[@estimatr].

After checking some basic characteristics of the original data, I changed individual score variables into numeric form for further analysis. Then, I filtered observations that should not be included since this evaluation program is for Apartment Building with three or more storey or 10 or more units to comply with building maintenance standards. Also, to study the building age problem in Toronto, I created a New Variable to show the building age when encountering the evaluation. Then, a single measurement is created or location as distance to downtown using CN Tower (latitude = 43.642567,-79.387054) as the benchmark.\

According to RentSafeTO, buildings must undergo evaluation at least once every three years. If buildings score 65 per cent or less, then the next evaluation will take place within one year. If the score is between 66 – 85 per cent, the next evaluation will take place within two years and if buildings score 86 percent and above the next evaluation with be within three years[@TO]. So I created a new variable that separate observations into groups by this score criteria. To see if Toronto housing really has a building age problem, I created a new variable that separate observations into groups by the building age when conducting the evaluation.\

There are three types of property included: private, Toronto Community Housing Corporation (TCHC) or social housing[@data]. Table 2 provides an overview of the cleaned dataset grouping by property type with changes mentioned above.



```{r, echo=FALSE, message=FALSE}

table <- datasummary_balance(formula = ~PROPERTY_TYPE,
                    data = data_cleaned,
                    title = "Descriptive Statistics for Results of Evaluation Grouping by Property Types",
                    notes = "Data Source: Open Data Toronto",
                    dinm=FALSE)
landscape(table, margin = NULL)
```



## 2.3 Data Characteristic

First of all, we can see that most of the buildings included for the evaluations are private property from Table 3. However, in Table 2, the average number of areas evaluated is 17 which is the lowest one among three property types. This implies that the current criteria of evaluation lack of certain measurements for the majority type of the apartment buildings. This could be adjusted by modifying evaluation items to better incorporates specific factors of private buildings. In addition, Table 3 shows that approximately 90% of the buildings could not score higher than 85%. The overall condition is not satisfied for building a safe environment. Most apartment buildings need close attention of safety monitoring.

 

```{r, echo=FALSE}
data_cleaned %>%
  datasummary_skim(type = "categorical",
                   title = "Summary of Property Type and Score Group under RentSafeTO for Evaluation")
```





```{r, echo=FALSE}
data_cleaned %>% 
  ggplot(mapping = aes(x = Score_Group, fill = PROPERTY_TYPE)) +
  labs(title = "Toronto Apartment Building Evaluation Score by Property Type",
        x = "Score Group",
        y = "Count") +
  theme_minimal() +
  geom_bar()


```
\begin{center} 

Figure 1: Toronto Apartment Building Evaluation Score Distribution by Property Type

\end{center}

To examine whether different property type would lead to biased evaluation, Figure 1 is generated to demonstrate the proportion of each property type under each score group. The main priority of Social Housing is to be more affordable than private renting[@social]. So, the benefits of renting from a housing association or council usually forms a stereotype that the building environment would be relatively worse than private property. However, combining the observations from Table 2, the average score of social housings does not appear to be significantly lower than private buildings. Instead, only the apartment buildings provided by TCHC would need improvements in building safety.

```{r, echo=FALSE}
data_cleaned %>% 
  ggplot(mapping = aes(x = Building_Age_Group)) + 
   labs(title = "Toronto Apartment Building Age Distribution",
        x = "Building Age Group",
        y = "Count") +
  theme_minimal() +
  geom_bar(fill = "blue", color = "black")


```

\begin{center} 

Figure 2: Toronto Apartment Building Age Distribution

\end{center}

To further analyze the building age problem, Figure 2 is created to present the age distribution of the buildings. It shows that the majority of the buildings has been 30-90 years old which could face severe risks of constructure deterioration. The significant proportion of aged buildings in Toronto could derive from the rapid urbanization. However, we could see from Table 2 that none of the evaluated item is associated with the risk of aging.

```{r, echo=FALSE}
data_cleaned%>%
  ggplot(mapping = aes(x = SCORE, y = Building_Age_at_Evaluation, color = PROPERTY_TYPE)) +
  labs(title = "Relationship between Score and Building Age of Different Property Type ",
        x = "Score Group",
        y = "Building Age at Evaluation") +
  theme_minimal() +
  geom_point()
```


\begin{center} 

Figure 3: Scatterplot for Relationship between Score and Building Age at Evaluation of Different Property Type

\end{center}

Referring to Figure 3, most private buildings has aged over 50 years while most social housing are built less than 50 years. Combining with Table 2, it implies that there could be a potential negative relationship between safety evaluation score and building age. As mentioned above, the average score of social housing is even relatively higher than that of private properties. To further investigate the relationship, we would perform a Simple Linear Regression to see if this could explain a significant portion of relationship between safety score and building age. The dependent variable in this case would be the evaluation score, and the independent variable is our building age. 

Figure 4 demonstrate the scatterplot of the relation between evaluation score and building age with a fitted regression line. The fitted regression line is downward sloping implying that the relation between safety score and building age should be negative as we expected. However, we could see that the fitted regression line does not has a large absolute value of slope which means that the relation is not very strong. On the other hand, even though the fitted regression line does not appear to be very steep, building age should still be considered as an essential part when evaluating building safety. 
```{r, echo=FALSE, message=FALSE}
ggplot(data_cleaned, aes(x=Building_Age_at_Evaluation, y=SCORE))+
  theme_minimal()+
geom_point() +
geom_smooth(method = lm, se = FALSE)+
labs( x = "Building Age at Evaluation", y ="SCORE" )
```
\begin{center} 

Figure 4: Scatterplot with Fitted Regression Line between Score and Building Age

\end{center}


# 3 Limitations and Conclusion
One of the limitations of the analysis conducted is that Condo buildings, townhomes, or units in a private home (basement or main floor apartment) are not part of the RentSafeTO program[@TO]. This would imply that our results is not appropriate for generalization to a larger population. Also, the relation we discovered between the safety score and building age may not be appropriate as Linear regression may not be the best model to fit. 

Under current regulation system, buildings that score 50 per cent or less on the evaluation must undergo a comprehensive inspection and audit of all common areas. After the audit, investigations are opened for all service requests received. In addition, details about the most recent building evaluation must be posted on the Tenant Notification Board. They must also be shared with all prospective tenants[@TO]. 

However, to better provoke the awareness of housing safety, it is necessary for the government to develop a more completed evaluation system that incorporates risks associated with building age. Through aging, any facility may experience progressive growth of risk, due to deterioration of building materials or other natural physical forces. In addition, the coverage of the evaluation program is very limited. RentSafeTO should attempt to expand the coverage while diversify evaluation criteria for different property types. United criteria is not appropriate for every property type which may cause errors in evaluation. Impacts from architecture structures should also be analyzed to see if it should be considered when evaluating building safety as next step.


\newpage
# Reference



